---
title: "Mileage and Transmission"
author: "Aseem Rohatgi"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, fig.align="center", fig.width=4, fig.height=4)
library(ggplot2)
library(dplyr)
```

# Overview

We at *Motor Trend* looked at a collection of cars and explored the relationship between a set of variables describing the car and miles per gallon that the car achieved. Particularly, we were interested in answering the following two questions:
+ “Is an automatic or manual transmission better for MPG”
+ "Quantify the MPG difference between automatic and manual transmissions"

We found that 
+
+

# Data

We used the data we collected on 32 automobiles('73 & '74 models). We have stored the data in a data frame called `mtcars`. Let's look at a summary of it.

```{r summary}
summary(cbind(mtcars[,-(8:9)],
              vs=as.factor(mtcars[,8]),
              am=as.factor(mtcars[,9])))
```

We are interested in the how MPG, stored in the variable `mpg`, change with the transmission type, stored in the variable `am` as 0 for automatic, and 1 for manual. Let's look at a boxplot for the data.

```{r boxplot, fig.width=3, fig.height=3}
ggplot(mtcars) + geom_boxplot(aes(x=factor(am), y=mpg, fill=factor(am))) +
  labs(x="Transmission", fill="am", y="MPG") + theme(legend.position="none")
```

This initial look at the data suggests that there is an obvious relationship between the transmission type of the automobile and miles per gallon that it achieves. We performed regression analysis to solidify our intuition with statistical significance.

# 
